Exploring Potential 6LoWPAN Traffic Side Channels

### 2017-04-06 Writing/iotprotocols.tex
Update.

\section{A Typical IoT Protocol Stack}
%A typical sensor network is built as described in \ref{Protocols}.

\begin{table}[!h]
	\centering
	\input{tbl/protocols.tex}
	\caption{Protocol stack for sensor networks. (* are optional.)\label{Protocols}}
\end{table}

@@

\subsection{Our experimental network}
%Say something about devices, and choice of OS, etc.

>>>>

\section{A Typical IoT Protocol Stack}
%A typical sensor network is built as described in \ref{Protocols}.
%IoT Protocols
Many protocols are proposed for different IoT applications adapting to various requirements. For example, some smart houses simply use WiFi for connectivity and VANETs\footnote{Vehicular ad hoc networks} may adopt DSRC\cite{DSRC}.

%WSN Protocols
In this paper we focused on 6LoWPAN\cite{rfc4944} which is based on 802.15.4\cite{802154}. These standards are designated for constrained environments such as Wireless Sensor Networks,  but other competing standards exist at different layers. Bluetooth Low Energy(BLE)\cite{BLE} is a strong competitor to 802.15.4 as well as the LiFi\cite{LiFi} technology. Zigbee\cite{Zigbee} is originally intended as a collective protocol over 802.15.4 but it has been recently adapted for IP-based network as well in \cite{ZigbeeIp} in order to compete with 6LoWPAN. The RIME stack\cite{RIME} proposed a set of non-layered primitive over 802.15.4 but it is likely to be phased-out probably due to the reluctance of interoperability with TCP/IP protocol stack of Internet.

%6LoWPAN OS
6LoWPAN is supported by several competing IoT Operating Systems, including Contiki OS\cite{Contiki}, OpenWSN\cite{OpenWSN}, FreeRTOS\cite{FreeRTOS} and the latest RIOT\cite{RIOT}. Contiki OS is chosen for our experiments for its customisability.

\subsection{Our experimental network}
%Say something about devices, and choice of OS, etc.
%Hardware
Our experiment networks are constructed using hardware of TelosB\cite{TelosB} and CC2538\cite{CC2538}. TelosB is a low cost sensor powered by MSP430 with AES co-processor representing the low-end devices. CC2538 is the high end device powered by ARM Cortex-M3 with multiple cryptographic processors including AES, RSA, SHA-2 and ECC, suggesting for security usage.
%Should I cite our previous paper...?

%Software
On the software aspect, we by principle adopted the default settings of Contiki OS, except enabling the 802.15.4 Security upon its study. For Layer 4\cite{OSI} and above protocols, we considered the widely accepted combination of CoAP over DTLS(optional) over UDP\footnote{CoAPs is equivalent to CoAP over DTLS.}. \Cref{Protocols} summarises our choice of protocol stack in a simplified OSI\cite{OSI} model.

\begin{table}[!h]
	\centering
	\input{tbl/protocols.tex}
	\caption{Protocol stack for our experiments(* is optinal)\label{Protocols}}
\end{table}

@@

%Avaible security options.

### 2017-04-06 Writing/relatedwork.tex
Update.

Reflecting to IoT applications, most of the attacks may still be applicable exploiting the same features of packet length and response time, as we intend to demonstrate in this paper. Considering the future vision that IoT devices could be connected to the Internet with even more sensitive data flowing over different networks, the task of designing secure IoT applications poses to be ever challenging.

>>>>

Reflecting to IoT applications, most of the attacks may still be applicable exploiting the same features of packet length and response time, as we intend to demonstrate in this paper. Considering the future vision that IoT devices could be connected to the Internet with even more sensitive data flowing over different networks (imaging to control your door lock from a web browser on your desktop), the task of designing secure IoT applications poses to be ever challenging.

### 2017-04-06 Writing/submission.tex
Update.

\usepackage{graphicx}
\usepackage{graphics}
%\usepackage{subfigure}
\usepackage{cite}
\usepackage{url}
\usepackage{subcaption}
\usepackage{multirow}
% for alternatives, see appendix of manual

@@

\title{Exploring Potential 6LoWPAN Traffic Side Channels}
\author{Yan Yan \& Elisabeth Oswald \& Theo Tryfonas}

@@

\input{relatedwork}
%\subsection{Related Work}
%Write about attacks that exploit packet length and response times in the context of TCP/IP. Reflect on web applications and how the same might happen if more interesting applications might be deployed that involve IoT objects.

@@

\subsection{Guidance to IoT software developers}
%{\it
%Pad to MTU is recommended as it completely hides the length of plaintext. This method only induces slight overhead as 6LoWPAN MTU is small and requires only negligible computation.
%
%Unifying the drivers by strictly apply the standard MTU. Applications should be compatible to the standard anyway. 
%}
A general countermeasure to avoid leaks related to the packet length leakage is padding, as studied in \cite{Peekaboo}. Whilst padding to MTU is considered inefficient for Internet, it is in fact highly appropriate for 6LoWPAN because:
\begin{itemize}
	\item It completely hides the length of original plaintext.
	\item 6LoWPAN has only a low MTU of 127 bytes; therefore the overhead is acceptable.
	\item It induces negligible computational overhead.
\end{itemize}

@@

\subsection{Guidance for IoT software developers}
%Two options:
%\begin{itemize}
%	\item Randomise response time: needs to be resilient to statistical analysis.
%	\item Threashold response time (discard or reply in a constant time): performance tradeoff. Recommended for unreliable transmissions.
%\end{itemize}
It is notoriously difficult to produce time-constant code. But two approaches are available to a software developer:
\begin{itemize}
	\item Randomly delay the response. This essentially adds noise to the measurements of the adversary.

@@

\appendix

And this is my Appendix.

\subsection*{Appendix Subsection}

Some text.

>>>>

\usepackage{graphicx}
\usepackage{graphics}
%\usepackage{subfigure}
\usepackage{cite}
\usepackage{url}
\usepackage{subcaption}
\usepackage{multirow}

\input{packages}

% for alternatives, see appendix of manual

@@

\title{Exploring Potential 6LoWPAN Traffic Side Channels}
\author{Yan Yan \and Elisabeth Oswald \and Theo Tryfonas}

@@

\subsection{Related Work}
%Write about attacks that exploit packet length and response times in the context of TCP/IP. Reflect on web applications and how the same might happen if more interesting applications might be deployed that involve IoT objects.

%Literatures about Traffic Analysis...
Traffic Analysis is well studied in the context of encrypted Internet traffic, especially for web applications based on HTTPs and TCP/IP. The landmark study by Chen et al. \cite{WebSidechannel} discussed different side channel attacks against web applications and \cite{SuggestBox} studied the practicability of an attack specifically targeted Google and Bing search boxes. Later work by Mather and Oswald \cite{PinpointWeb} proposed the use of Mutual Information to pinpoint the potential leakage points in web traffic. For non-HTTPs applications, the papers \cite{AppleMsg}, \cite{Language} and \cite{VideoTraffic} described attacks against encrypted text, voice and video traffic respectively. Machine learning is widely used to analyse the traffic, and behaviours of different classifiers are studied by \cite{HClassifier} and \cite{Peekaboo}. Based on all these published works we can conclude that two features, the packet length and response time, are the most exploited ones among all attacks. Different countermeasures were studied by \cite{TrafficMorphing}, \cite{HTTPOS} and \cite{FTE}.

Reflecting on IoT applications, we stipulate that most of these attacks may still be applicable, as we intend to demonstrate in this paper. Considering the future vision that IoT devices could be indeed connected to the Internet with even more sensitive data flowing over different networks, the task of designing secure IoT applications becomes increasingly challenging.

%Literatures about 6LoWPAN security
With regard to the aspect of protocol design, the recent paper \cite{6LoWPANAtk} summarised some known flaws of 6LoWPAN, including its susceptibility to the Fragmentation Attack\cite{FragAtk}, Sinkhole Attack\cite{Sinkhole}, Hello Flood Attack\cite{HelloFlood}, Wormhole Attack\cite{Wormhole} and Blackhole Attack\cite{Blackhole}. In addition, \cite{802154SecIssues} reported certain problematic designs in 802.15.4 security\cite{802154}.

TODO: if space and time say some more about these attacks. At least say if they relate to the type of leakage we are exploiting or if they are completely independent (which is what I would guess).

@@

@@

In this paper we explore, for the first time, the use of packet lengths and response times, which are protocol level side channels, as means to recover information about IoT `things'. We do this via some experiments, which we base on two extremely popular devices running on a popular open source OS, with a typical stack of protocols. Whilst we do not cover a wide range of devices, the fact that two of the most popular devices show the characteristics that we hypothesise, gives credibility to our results. Our results show that it is possible (in principle) to recover information about a device and its function (i.e. the hardware and the software that runs on it) via inspecting encrypted traffic that it produces. We also point out that ICMP messages can be distinguished from each other despite the use of encryption.

In order to mitigate the leakage that is given by packet lengths, previous works recommends padding \cite{Peekaboo}. We echo this recommendation. Whilst padding to MTU is considered inefficient for the Internet, it is in fact highly appropriate for 6LoWPAN because:
\begin{itemize}
	\item It completely hides the length of original plaintext.
	\item 6LoWPAN has only a low MTU of 127 bytes; therefore the overhead is acceptable.
	\item It induces negligible computational overhead.
\end{itemize}

%Two options:
%\begin{itemize}
%	\item Randomise response time: needs to be resilient to statistical analysis.
%	\item Threashold response time (discard or reply in a constant time): performance tradeoff. Recommended for unreliable transmissions.
%\end{itemize}
In order to mitigate the leakage given by response times, the natural countermeasure is to write time-constant code, which is known to be notoriously difficult. But two approaches are available to a software developer:
\begin{itemize}
	\item Randomly delay the response. This essentially adds noise to the measurements of the adversary.

@@

%\appendix

%And this is my Appendix.

%\subsection*{Appendix Subsection}

%Some text.
